---
output: html_document
---
## Download the file to the data folder
```{r}
if(!file.exists("./data")){dir.create("./data")}
fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
download.file(fileUrl,destfile="./data/Dataset.zip",method="curl")
```

## Unzip the file
```{r}
unzip(zipfile="./data/Dataset.zip",exdir="./data")
```

##Get the list of the files in the directory UCI HAR Dataset
```{r}
path_rf <- file.path("./data" , "UCI HAR Dataset")
files<-list.files(path_rf, recursive=TRUE)
files
```

##Reading the Activity files
```{r}
dataActivityTest  <- read.table(file.path(path_rf, "test" , "Y_test.txt" ),header = FALSE)
dataActivityTrain <- read.table(file.path(path_rf, "train", "Y_train.txt"),header = FALSE)
```

## Reading the Subject files
```{r}

dataSubjectTrain <- read.table(file.path(path_rf, "train", "subject_train.txt"),header = FALSE)
dataSubjectTest  <- read.table(file.path(path_rf, "test" , "subject_test.txt"),header = FALSE)
```
## Reading the Features files
```{r}
dataFeaturesTest  <- read.table(file.path(path_rf, "test" , "X_test.txt" ),header = FALSE)
dataFeaturesTrain <- read.table(file.path(path_rf, "train", "X_train.txt"),header = FALSE)
```
## Look at the properties of the 3 files
```{r}
str(dataActivityTest)
str(dataActivityTrain)
str(dataSubjectTrain)
str(dataSubjectTest)
str(dataFeaturesTest)
str(dataFeaturesTrain)
```

## Merging the training and test data sets
## Rbind data tables
```{r}
dataSubject <- rbind(dataSubjectTrain, dataSubjectTest)
dataActivity<- rbind(dataActivityTrain, dataActivityTest)
dataFeatures<- rbind(dataFeaturesTrain, dataFeaturesTest)
```

## Naming of variables
```{r}
names(dataSubject)<-c("subject")
names(dataActivity)<- c("activity")
dataFeaturesNames <- read.table(file.path(path_rf, "features.txt"),head=FALSE)
names(dataFeatures)<- dataFeaturesNames$V2
```

## Merging columns to get dataframe "data"
```{r}
dataCombine <- cbind(dataSubject, dataActivity)
Data <- cbind(dataFeatures, dataCombine)
```

## extracting mean and Std.Dev for each measurement
```{r}
subdataFeaturesNames<-dataFeaturesNames$V2[grep("mean\\(\\)|std\\(\\)", dataFeaturesNames$V2)]
```

## creating subsets of selected variables
```{r}
selectedNames<-c(as.character(subdataFeaturesNames), "subject", "activity" )
Data<-subset(Data,select=selectedNames)
str(Data)
```

## Creating descriptive names for activities in data set
```{r}
activityLabels <- read.table(file.path(path_rf, "activity_labels.txt"),header = FALSE)
head(Data$activity,30)
```

## creating appropriate labels 
```{r}
names(Data)<-gsub("^t", "time", names(Data))
names(Data)<-gsub("^f", "frequency", names(Data))
names(Data)<-gsub("Acc", "Accelerometer", names(Data))
names(Data)<-gsub("Gyro", "Gyroscope", names(Data))
names(Data)<-gsub("Mag", "Magnitude", names(Data))
names(Data)<-gsub("BodyBody", "Body", names(Data))
names(Data)
```

## Creating cleaned data set and output
```{r}
library(plyr);
Data2<-aggregate(. ~subject + activity, Data, mean)
Data2<-Data2[order(Data2$subject,Data2$activity),]
write.table(Data2, file = "tidydata.txt",row.name=FALSE)
```


